When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-05-27 doc/paper/body.tex
proofread, about to submit to arxiv

The theory for GRA suffers from similar, but much more severe,
limitations as for CRO, WSA, and SAV. GRA observations are
characterized by a consistent trend back towards negative ET response
at higher VPD, which the functional form of our theory is incapable of
accounting for. As compared to CRO, WSA and SAV, the divergence
between the theory and the observations is far greater for GRA,
biasing 40-50\% of observations at higher VPD. In addition for the
potential for soil moisture to alter $uWUE$, there are other sources
of plant heterogeneity specific to GRA (and some extent CRO) that may
alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
Methods section (Section \ref{methods}). We do not account for
variability in plant height and surface roughness, or differences in
C3 vs C4 photosynthesis and water strategies, which we might expect to
vary substantially across sites, years, and season for GRA. These
deficiencies could largely explain the inability of our theory to
exactly match the observations in croplands. For example, a
superposition of sites with C3 plants (at low environmental VPD) and
C4 plants (at high environmental VPD) would explain the observed shift
back towards negative ET response at high VPD when all sites are
binned together, as in Figure \ref{test_sign}. We hypothesize that the
theory would validate against observations much better if these
sources of variability were accounted for, at a cost of increased
complexity and analytical opacity.

@@

Indeed, for all PFTs, there is some slight dependence of
$\sigma \cdot uWUE$ on SWC, especially at low SWC (Figure
\ref{swc_boxplot}). The portion of observations for which our theory
is biased by SWC-limitations varies by PFT, due to the nonlinear
threshold at which soil moisture availability limits plant
function. For CRO, MF and DBF, soil water content only matters for
about the lowest 5\% of observations (each box is 5\% of observations
in Figure \ref{swc_boxplot}). And, CRO, relative to DBF, has a very
weak dependence on soil moisture, which is reflective of the high
likelihood that CRO sites are optimally irrigated and not water
stressed, suggesting that observed departures from the theory for CRO
are due to other factors (i.e. different photosynthesis pathways; C3/C4) than soil
moisture. For ENF, EBF, SAV, and WSA, systematic SWC-induced biases in
$\sigma \cdot uWUE$ emerge in about the lowest 20-30\% of SWC
conditions, although large variability in the SWC-$(\sigma \cdot uWUE)$
relationship hampers interpretation for WSA. CSH presents a special
case: the limited number of sites (2) preclude comment on the
exact PFT-wide relationship.

>>>>

The theory for GRA suffers from similar, but much more severe,
limitations as for CRO, WSA, and SAV. GRA observations are
characterized by a consistent trend back towards negative ET response
at higher VPD, which the functional form of our theory is incapable of
accounting for. As compared to CRO, WSA and SAV, the divergence
between the theory and the observations is far greater for GRA,
biasing 40-50\% of observations at higher VPD. In addition for the
potential for soil moisture to alter $uWUE$, there are other sources
of plant heterogeneity specific to GRA (and some extent CRO) that may
alter $uWUE$ (or $g_1$) or invalidate other assumptions made in the
methods section (Section \ref{methods}). We do not account for
variability in plant height and surface roughness, or differences in
C3 vs C4 photosynthesis and water strategies, which we might expect to
vary substantially across sites, years, and season for GRA. These
deficiencies could largely explain the inability of our theory to
exactly match the observations in croplands. For example, a
superposition of sites with C3 plants (at low environmental VPD) and
C4 plants (at high environmental VPD) would explain the observed shift
back towards negative ET response at high VPD when all sites are
binned together, as in Figure \ref{test_sign}. We hypothesize that the
theory would validate against observations much better if these
sources of variability were accounted for, at a cost of increased
complexity and analytical opacity.

@@

Indeed, for all PFTs, there is some slight dependence of $\sigma \cdot
uWUE$ on SWC, especially at low SWC (Figure \ref{swc_boxplot}). The
portion of observations for which our theory is biased by
SWC-limitations varies by PFT, due to the nonlinear threshold at which
soil moisture availability limits plant function. For CRO, MF and DBF,
soil water content only matters for about the lowest 5\% of observations
(each box is 5\% of observations in Figure \ref{swc_boxplot}). And, CRO,
relative to DBF, has a very weak dependence on soil moisture, which is
reflective of the high likelihood that CRO sites are optimally irrigated
and not water stressed, suggesting that observed departures from the
theory for CRO are due to other factors (e.g. different photosynthesis
pathways; C3/C4) than soil moisture. For ENF, EBF, SAV, and WSA,
systematic SWC-induced biases in $\sigma \cdot uWUE$ emerge in about the
lowest 20-30\% of SWC conditions, although large variability in the
SWC-$(\sigma \cdot uWUE)$ relationship hampers interpretation for
WSA. CSH presents a special case: the limited number of sites (2)
preclude comment on the exact PFT-wide relationship.
